On Policy Gradients

### 2019-08-07 paper.tex
Convert to IEEE article layout

\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
\RequirePackage{fix-cm}
\documentclass[final]{svjour3}
\smartqed
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{ {./images/} }
\usepackage{amsmath,amssymb}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{environ}
\usetikzlibrary{shapes,arrows,positioning,calc}
\begin{document}

\title{Policy Gradient Algorithms}
\subtitle{An Introduction}

\author {Mattis Manfred K{\"a}mmerer}

\institute{TU Darmstadt}

\date{Received: date / Accepted: date}

@@

The goal of policy gradient approaches is to find a policy in a given class of policies which maximizes the expected return.
Given a differentiable model of the policy, we want to apply a gradient-ascent technique to reach the local optimum. 
We mainly use gradient ascent, because it is theoretically well researched.
The main issue is that the policy gradient with respect to the expected return is not available, thus we need to estimate it.
As they also tend to require on-policy data for the gradient estimate, the biggest weakness of policy gradient algorithms is sample efficiency.
For this reason, most research is focused on finding algorithms with improved sample efficiency.
This paper provides a deep introduction to policy gradient that shows the development of policy gradient approaches, and should enable the reader to follow current research in the area.

\keywords{reinforcement learning \and policy gradients \and actor-critic \and natural gradient }
\end{abstract}

@@

\paragraph{Definitions.} 
We define states $s \in \mathbb{S}$, actions $a \in \mathbb{A}$, and rewards $r \in \mathbb{R}$. 
A trajectory $\tau := (s_0, a_0, $ $s_1, a_1, \dots, s_T, a_T)$ is generated by drawing $s_0 \sim \mu_0(s_0)$ according to the distribution over initial states $\mu_0(s_0)$, and successively sampling $a_t \sim \pi(a_t|s_t)$ according to the policy $\pi$ parameterized by $\theta$, and $s_{t+1} \sim p(s_{t+1}|s_t,a_t)$ until the horizon $T$, or a terminal state is reached. 
At each time step, we receive a reward according to $r_t = r(s_t, a_t) \equiv \mathbb{E}_{s'}\left[r(s_t,a_t,s')\right]$. 
A trajectory can also be called roll-out or episode, though the term episode implies it ends in a terminal state.
We assume a Markov Decision Process (MDP), meaning the probability distribution of the next states is independent of past states $s_{0:t-1}$ given the present state $s_t$ and action $a_t$, 
\begin{equation}
	p(s_{t+1}|s_t,a_t)=p(s_{t+1}|s_{0:t},a_{0:t}).
\end{equation}
Where we define $i:j$ with $i,j \in \mathbb{N}, i < j$ as an index over all integers from $i$ to $j$, i.e., $s_{i:j} \equiv s_i, s_{i+1}, \dots, s_j$. 
We assume no additional prior knowledge about the environment, meaning we assume the probability of a trajectory is 
\begin{equation}
	p_\pi(\tau) = \mu_0(s_0) \prod_{t=0}^{T-1} p(s_{t+1}|s_t, a_t) \pi(a_t|s_t).
\end{equation}

@@

\paragraph{Policy gradient.} 
Our goal with respect to episodes is to maximize the expectation of the total reward, also called expected return. 
The total reward in the horizon $T$ is $\sum_{t=0}^{T} r_{t}$. 
We additionally introduce a discount factor $\gamma \in [0,1)$.
Intuitively, this reflects the idea that the relevance of later actions declines, and ensures that the return is finite, even for the infinite horizon $T \to \infty$. 
The discounted total reward is 
\begin{equation}
  \mathcal{R}^\tau \equiv \mathcal{R}_0^T := \sum_{t=0}^{T} \gamma^t r_t.
  \label{eqn:acc-reward}
\end{equation}

@@

\paragraph{Finite-difference gradients.} 
A simple approach for gradient estimation is to choose a small $\delta\theta$, and evaluate the new policy given the slightly changed parameters as in 
\begin{equation}
	\nabla_\theta J(\theta) \approx \frac{J(\theta+\delta\theta)J(\theta-\delta\theta)}{2\delta\theta}.
\end{equation} 
This can lend a good estimate of the gradient given a small $\delta \theta$, and is generally called the symmetric derivative. 
However, finite-difference gradients suffer from the curse of dimensionality, and can require very small $\delta\theta$.
Thus, finite-difference gradients only work well in specific scenarios, but should not be discarded due to simplicity.

\paragraph{Value functions.} 
Given we know the actual value of a state, i.e. the expected return we will get starting from state $s_t$, this function can be used to evaluate the performance of our policy, and can be written as
\begin{equation}
	V^{\pi}(s_t) := \mathbb{E}_{\substack{s_{t+1:h} \\ a_{t:h}}}\left[\mathcal{R}_t^T\right].
	\label{eqn:v}
\end{equation}
In addition to the value function, we also define the state-action value function, often called Q-function. 
Instead of the expected accumulated reward starting from state $s_t$, this function gives the expected accumulated reward given an action $a_t$ is selected in state $s_t$, 
\begin{equation}
	Q^{\pi}(s_t, a_t) := \mathbb{E}_{\substack{s_{t+1:h} \\ a_{t+1:h}}}\left[\mathcal{R}_t^T\right].
	\label{eqn:q}
\end{equation}
As we will see, this function also gives us the true gradient of $J(\theta)$, though in general we need to estimate it. 
Using the value function, and the Q-function, we can derive a better estimate of the policy gradient.

\paragraph{Likelihood-ratio gradients.} 
For this derivation, we will change the perspective a bit, requiring some additional definitions. 
We define ${\mu_\pi}_i = \sum_{t=0}^{\infty} \gamma^t p(s_t=s_i | s_0, \pi)$ as the discounted state distribution, though it does not sum up to one without normalization, which can be achieved by multiplying by $(1-\gamma)$. 
Note that $\mu_\pi$ is equivalent to the discounted state visit count $d^\pi$ introduced by Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806}.
Further, we define $P_\pi$ as the transition matrix, i.e. ${P_\pi}_{i,j}=\sum\nolimits_k p(s_j|s_i,a_k)\pi(a_k|s_i)$, $r_\pi$ as the mean rewards for all states given by ${r_\pi}_i = \sum_j r(s_i,a_j)\pi(a_j|s_i)$, and $\mu_0 = \left[\mu_0(s_0),\mu_0(s_1),\ldots\right]^T$ as a vector representing the initial state distribution.
Finally, we define $V_\pi = \left[ V_\pi(s_0), V_\pi(s_1), \ldots \right]^T$, from which it follows that $V_\pi = \mu_\pi r_\pi$, so we can reformulate the problem as 
\begin{align}
	\underset{\theta}{\text{max}}\ J(\theta) &= \mu_0^T V_\pi \\
	\text{s.t.}\ V_\pi &= r_\pi + \gamma P_\pi V_\pi . \nonumber
\end{align}
Since $\mu_0^T$ does not depend on $\theta$, it is obvious that 
\begin{align*}
	\nabla_\theta J(\theta) &= \nabla_\theta \mu_0^T V_\pi = \mu_0^T \nabla_\theta V_\pi.
\end{align*}
We can replace $\nabla_\theta V_\pi$ using
\begin{align*}
	\nabla_\theta V_\pi &= \nabla_\theta \left( r_\pi + \gamma P_\pi V_\pi \right)
	\\ \nabla_\theta V_\pi &= \nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi + \gamma P_\pi \nabla_\theta V_\pi
	\\ (I-\gamma P_\pi) \nabla_\theta V_\pi &= \nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi
	\\ \nabla_\theta V_\pi &= (I - \gamma P_\pi)^{-1} (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi), 
\end{align*}
and find that
\begin{align*}
	\mu_\pi &= \mu_0 + \gamma P_\pi^T \mu_\pi \nonumber \\
	(I- \gamma P_\pi^T)\mu_\pi &= \mu_0 \nonumber \\
	\mu_\pi^T &= \mu_0^T (I- \gamma P_\pi)^{-1} ,
\end{align*}
which we can take back into the gradient equation 
\begin{align}
	\nabla_\theta J(\theta) &= \mu_0^T (I - \gamma P_\pi)^{-1} (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi) \nonumber
	\\ &= \mu_\pi^T (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi) \nonumber
	\\ &\equiv \sum\nolimits_{i,j} \mu(s_i) \nabla_\theta\pi(a_j|s_i) Q_\pi(s_i,a_j) \label{eqn:likelihood-equiv}
	\\ &= \sum\nolimits_{i,j} \mu(s_i) \pi(a_j|s_i) \nabla_\theta\log\pi(a_j|s_i) Q_\pi(s_i,a_j). \nonumber
\end{align}
The equivalence in \eqref{eqn:likelihood-equiv} comes from the observation that $\nabla_\theta \pi(a|s)$ is distributed over the addition $\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi$.
When we take out this common factor, $Q_\pi(s,a)$ remains.
Then, we used $\nabla_\theta \pi(a|s) = \pi(a|s)\nabla_\theta\log\pi(a|s)$, obtained from the likelihood ratio $\nabla_\theta \log p(x|\theta) = \frac{\nabla_\theta p(x|\theta)}{p(x|\theta)}$. 
This gives us the likelihood-ratio gradient
\begin{equation}
	\nabla_\theta J(\theta) = \mathbb{E}_{\substack{\ s \sim \mu_\pi \\a \sim \pi}} \Big[\nabla_\theta{\log\pi(a|s)}Q_\pi(s,a)\Big],
	\label{eqn:like-grad}
\end{equation}
intuitively meaning we should to increase the probability of actions that lead to higher Q-values.
This formulation enables us to calculate the gradient $\nabla_\theta J(\theta)$, while directly taking advantage of the MDP structure in the form of $Q_\pi(s,a)$.

@@

\paragraph{Episode-based updates.} 
A very general optimization approach to this optimization problem are episodic algorithms. 
We take a search distribution $p(\theta|\omega)$ over the parameter space of the policy class $\pi$, and sample acting policies from that distribution. 
The policy class $\pi$ is most often chosen deterministic. 
Using these policies, we sample trajectories $\tau$, and update the search policy using the returns of our sampled roll-outs
\begin{equation}
	\nabla_\theta J(\theta) \approx \sum_{t=0}^T \nabla_\omega \log p(\theta|\omega) \mathcal{R}_t^T.
\end{equation}
The resulting algorithms are black-box optimizers, and as such are largely applicable, but can not use any temporal information and have a lot of variance.
Given these insights, we require a way to design algorithms that improve the acting policy stepwise by observing each interaction with the environment.

\paragraph{Step-based updates.}
The first class of algorithms developed to update a policy directly using a critic are called REINFORCE \cite{Williams92simplestatistical}.
REINFORCE samples a complete episode, at which point we can calculate the actual state-action value by traversing backwards over the trajectory, and estimates
\begin{equation}
	\nabla_\theta J(\theta) \approx \nabla_\theta\log\pi(s) \left(Q(s_t,a_t) - b_\tau \right).
	\label{eqn:reinforce}
\end{equation}
This is sometimes also called Monte-Carlo gradient estimation. 
However, given $b_\tau = 0, r_t > 0, \forall t=0,\dots,h$, we can only increase action probabilities. 
Obviously, we normalize to ensure $\forall s \in \mathbb{S}: \int_\mathbb{A}{\pi(a|s)da} = 1$. 
This means actions can only become less probable in relation to other actions. 
We find that this introduces more variance when learning from samples \cite{Sutton:1999:PGM:3009657.3009806}, and by that defeats the purpose of why we thought of this approach in the first place. 
One way to counter the variance is to use an effective baseline $b_\tau$.
Peters et al. \cite{4867} find that an estimate of the optimal baseline can be calculated by
\begin{equation}
	b_\tau = \frac
		{\left\langle 
			\left(\sum_{t=0}^T \nabla_\theta \log\pi(a_t|s_t) \right)^2 \sum_{t'=0}^T a_{t'} r_{t'} 
		\right\rangle}
		{\left\langle
			\left(\sum_{t=0}^T \nabla_\theta \log\pi(a_t|s_t) \right)^2
		\right\rangle},
\end{equation}
which does not affect the unbiasedness of the estimate.

@@

\begin{figure}
  \includegraphics[width=\textwidth]{nat-grad-adv}
  \caption{An experiment showing where the natural gradient has a great advantage \cite{Peters_IICHR_2003}. }\label{fig:nat-grad-adv}
\end{figure}

Using the Fisher information matrix $F_\theta$, and the gradient estimate we discussed in section \ref{sec:pge} gives us the definition
\begin{equation}
  \widetilde{\nabla}_\theta J(\theta) := F^{-1}_\theta \nabla_\theta J(\theta)
  \label{eqn:nat-grad}
\end{equation}
of the natural gradient. 
The Fisher information matrix represents the certainty we have on our estimate of the gradient and is defined as the covariance of the log likelihood function of a trajectory $\tau_{\pi}^T$, which as \cite{4863} show can be written as 
\begin{equation}
  F_\theta = \int_\mathbb{S} d^\pi(s) \int_\mathbb{A} \pi(a|s) \nabla_\theta \log{\pi(a|s)} \nabla_\theta \log{\pi(a|s)}^T dads.
  \label{eqn:F}
\end{equation}
Using a value function estimator and calculating the natural gradient, we get the natural policy gradient algorithm (NPG) \cite{NIPS2017_7233}.
But, if we recall the definition \eqref{eqn:like-grad} of likelihood-ratio gradients, and the compatible function approximator from \eqref{eqn:compatible-approximator}, we get
\begin{equation}
  \nabla_\theta J(\theta) = \int_\mathbb{S} {d^\pi(s) \int_\mathbb{A} \pi(a|s) \nabla_\theta \log \pi(a|s) {\nabla_\theta \log \pi(a|s)}^T dads}\ w \equiv F_\theta w.
  \label{eqn:J-equals-F}
\end{equation}

>>>>

\RequirePackage{fix-cm}
\documentclass[final]{IEEEtran}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{ {./images/} }
\usepackage{amsmath,amssymb}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{environ}
\usetikzlibrary{shapes,arrows,positioning,calc}

\begin{document}

\title{On Policy Gradients}

\author {Mattis Manfred K{\"a}mmerer}
\date{Received: date / Accepted: date}

@@

The goal of policy gradient approaches is to find a policy in a given class of policies which maximizes the expected return.
Given a differentiable model of the policy, we want to apply a gradient-ascent technique to reach the local optimum. 
We mainly use gradient ascent, because it is theoretically well researched.
The main issue is that the policy gradient with respect to the expected return is not available, thus we need to estimate it.
As they also tend to require on-policy data for the gradient estimate, the biggest weakness of policy gradient algorithms is sample efficiency.
For this reason, most research is focused on finding algorithms with improved sample efficiency.
This paper provides a deep introduction to policy gradient that shows the development of policy gradient approaches, and should enable the reader to follow current research in the area.
\end{abstract}

@@

\paragraph{Definitions} 
We define states $s \in \mathbb{S}$, actions $a \in \mathbb{A}$, and rewards $r \in \mathbb{R}$. 
A trajectory $\tau := (s_0, a_0, $ $s_1, a_1, \dots, s_T, a_T)$ is generated by drawing $s_0 \sim \mu_0(s_0)$ according to the distribution over initial states $\mu_0(s_0)$, and successively sampling $a_t \sim \pi(a_t|s_t)$ according to the policy $\pi$ parameterized by $\theta$, and $s_{t+1} \sim p(s_{t+1}|s_t,a_t)$ until the horizon $T$, or a terminal state is reached. 
At each time step, we receive a reward according to $r_t = r(s_t, a_t) \equiv \mathbb{E}_{s'}\left[r(s_t,a_t,s')\right]$. 
A trajectory can also be called roll-out or episode, though the term episode implies it ends in a terminal state.
We assume a Markov Decision Process (MDP), meaning the probability distribution of the next states is independent of past states $s_{0:t-1}$ given the present state $s_t$ and action $a_t$, 
\begin{equation}
	p(s_{t+1}|s_t,a_t)=p(s_{t+1}|s_{0:t},a_{0:t}).
\end{equation}
Where we define $i:j$ with $i,j \in \mathbb{N}, i < j$ as an index over all integers from $i$ to $j$, i.e., $s_{i:j} \equiv s_i, s_{i+1}, \dots, s_j$. 
We assume no additional prior knowledge about the environment, meaning we assume the probability of a trajectory is 
\begin{equation}
	p_\pi(\tau) = \mu_0(s_0) \prod_{t=0}^{T-1} p(s_{t+1}|s_t, a_t) \pi(a_t|s_t).
\end{equation}

@@

\paragraph{Policy gradient} 
Our goal with respect to episodes is to maximize the expectation of the total reward, also called expected return. 
The total reward in the horizon $T$ is $\sum_{t=0}^{T} r_{t}$. 
We additionally introduce a discount factor $\gamma \in [0,1)$.
Intuitively, this reflects the idea that the relevance of later actions declines, and ensures that the return is finite, even for the infinite horizon $T \to \infty$. 
The discounted total reward is 
\begin{equation}
  \mathcal{R}^\tau \equiv \mathcal{R}_0^T := \sum_{t=0}^{T} \gamma^t r_t.
  \label{eqn:acc-reward}
\end{equation}

@@

\paragraph{Finite-difference gradients} 
A simple approach for gradient estimation is to choose a small $\delta\theta$, and evaluate the new policy given the slightly changed parameters as in 
\begin{equation}
	\nabla_\theta J(\theta) \approx \frac{J(\theta+\delta\theta)J(\theta-\delta\theta)}{2\delta\theta}.
\end{equation} 
This can lend a good estimate of the gradient given a small $\delta \theta$, and is generally called the symmetric derivative. 
However, finite-difference gradients suffer from the curse of dimensionality, and can require very small $\delta\theta$.
Thus, finite-difference gradients only work well in specific scenarios, but should not be discarded due to simplicity.

\paragraph{Value functions} 
Given we know the actual value of a state, i.e. the expected return we will get starting from state $s_t$, this function can be used to evaluate the performance of our policy, and can be written as
\begin{equation}
	V^{\pi}(s_t) := \mathbb{E}_{\substack{s_{t+1:h} \\ a_{t:h}}}\left[\mathcal{R}_t^T\right].
	\label{eqn:v}
\end{equation}
In addition to the value function, we also define the state-action value function, often called Q-function. 
Instead of the expected accumulated reward starting from state $s_t$, this function gives the expected accumulated reward given an action $a_t$ is selected in state $s_t$, 
\begin{equation}
	Q^{\pi}(s_t, a_t) := \mathbb{E}_{\substack{s_{t+1:h} \\ a_{t+1:h}}}\left[\mathcal{R}_t^T\right].
	\label{eqn:q}
\end{equation}
As we will see, this function also gives us the true gradient of $J(\theta)$, though in general we need to estimate it. 
Using the value function, and the Q-function, we can derive a better estimate of the policy gradient.

\paragraph{Likelihood-ratio gradients} 
For this derivation, we will change the perspective a bit, requiring some additional definitions. 
We define ${\mu_\pi}_i = \sum_{t=0}^{\infty} \gamma^t p(s_t=s_i | s_0, \pi)$ as the discounted state distribution, though it does not sum up to one without normalization, which can be achieved by multiplying by $(1-\gamma)$. 
Note that $\mu_\pi$ is equivalent to the discounted state visit count $d^\pi$ introduced by Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806}.
Further, we define $P_\pi$ as the transition matrix, i.e. ${P_\pi}_{i,j}=\sum\nolimits_k p(s_j|s_i,a_k)\pi(a_k|s_i)$, $r_\pi$ as the mean rewards for all states given by ${r_\pi}_i = \sum_j r(s_i,a_j)\pi(a_j|s_i)$, and $\mu_0 = \left[\mu_0(s_0),\mu_0(s_1),\ldots\right]^T$ as a vector representing the initial state distribution.
Finally, we define $V_\pi = \left[ V_\pi(s_0), V_\pi(s_1), \ldots \right]^T$, from which it follows that $V_\pi = \mu_\pi r_\pi$, so we can reformulate the problem as 
\begin{align}
	\underset{\theta}{\text{max}}\ J(\theta) &= \mu_0^T V_\pi \\
	\text{s.t.}\ V_\pi &= r_\pi + \gamma P_\pi V_\pi . \nonumber
\end{align}
Since $\mu_0^T$ does not depend on $\theta$, it is obvious that 
\begin{align*}
	\nabla_\theta J(\theta) &= \nabla_\theta \mu_0^T V_\pi = \mu_0^T \nabla_\theta V_\pi.
\end{align*}
We can replace $\nabla_\theta V_\pi$ using
\begin{align*}
	\nabla_\theta V_\pi &= \nabla_\theta \left( r_\pi + \gamma P_\pi V_\pi \right)
	\\ \nabla_\theta V_\pi &= \nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi + \gamma P_\pi \nabla_\theta V_\pi
	\\ (I-\gamma P_\pi) \nabla_\theta V_\pi &= \nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi
	\\ \nabla_\theta V_\pi &= (I - \gamma P_\pi)^{-1} (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi), 
\end{align*}
and find that
\begin{align*}
	\mu_\pi &= \mu_0 + \gamma P_\pi^T \mu_\pi \nonumber \\
	(I- \gamma P_\pi^T)\mu_\pi &= \mu_0 \nonumber \\
	\mu_\pi^T &= \mu_0^T (I- \gamma P_\pi)^{-1} ,
\end{align*}
which we can take back into the gradient equation 
\begin{align}
	\nabla_\theta J(\theta) &= \mu_0^T (I - \gamma P_\pi)^{-1} (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi) \nonumber
	\\ &= \mu_\pi^T (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi) \nonumber
	\\ &\equiv \sum\nolimits_{i,j} \mu(s_i) \nabla_\theta\pi(a_j|s_i) Q_\pi(s_i,a_j) \label{eqn:likelihood-equiv}
	\\ &= \sum\nolimits_{i,j} \mu(s_i) \pi(a_j|s_i) \nabla_\theta\log\pi(a_j|s_i) Q_\pi(s_i,a_j). \nonumber
\end{align}
The equivalence in \eqref{eqn:likelihood-equiv} comes from the observation that $\nabla_\theta \pi(a|s)$ is distributed over the addition $\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi$.
When we take out this common factor, $Q_\pi(s,a)$ remains.
Then, we used $\nabla_\theta \pi(a|s) = \pi(a|s)\nabla_\theta\log\pi(a|s)$, obtained from the likelihood ratio $\nabla_\theta \log p(x|\theta) = \frac{\nabla_\theta p(x|\theta)}{p(x|\theta)}$. 
This gives us the likelihood-ratio gradient
\begin{equation}
	\nabla_\theta J(\theta) = \mathbb{E}_{\substack{\ s \sim \mu_\pi \\a \sim \pi}} \Big[\nabla_\theta{\log\pi(a|s)}Q_\pi(s,a)\Big],
	\label{eqn:like-grad}
\end{equation}
intuitively meaning we should to increase the probability of actions that lead to higher Q-values.
This formulation enables us to calculate the gradient $\nabla_\theta J(\theta)$, while directly taking advantage of the MDP structure in the form of $Q_\pi(s,a)$.

@@

\paragraph{Episode-based updates} 
A very general optimization approach to this optimization problem are episodic algorithms. 
We take a search distribution $p(\theta|\omega)$ over the parameter space of the policy class $\pi$, and sample acting policies from that distribution. 
The policy class $\pi$ is most often chosen deterministic. 
Using these policies, we sample trajectories $\tau$, and update the search policy using the returns of our sampled roll-outs
\begin{equation}
	\nabla_\theta J(\theta) \approx \sum_{t=0}^T \nabla_\omega \log p(\theta|\omega) \mathcal{R}_t^T.
\end{equation}
The resulting algorithms are black-box optimizers, and as such are largely applicable, but can not use any temporal information and have a lot of variance.
Given these insights, we require a way to design algorithms that improve the acting policy stepwise by observing each interaction with the environment.

\paragraph{Step-based updates}
The first class of algorithms developed to update a policy directly using a critic are called REINFORCE \cite{Williams92simplestatistical}.
REINFORCE samples a complete episode, at which point we can calculate the actual state-action value by traversing backwards over the trajectory, and estimates
\begin{equation}
	\nabla_\theta J(\theta) \approx \nabla_\theta\log\pi(s) \left(Q(s_t,a_t) - b_\tau \right).
	\label{eqn:reinforce}
\end{equation}
This is sometimes also called Monte-Carlo gradient estimation. 
However, given $b_\tau = 0, r_t > 0, \forall t=0,\dots,h$, we can only increase action probabilities. 
Obviously, we normalize to ensure $\forall s \in \mathbb{S}: \int_\mathbb{A}{\pi(a|s)da} = 1$. 
This means actions can only become less probable in relation to other actions. 
We find that this introduces more variance when learning from samples \cite{Sutton:1999:PGM:3009657.3009806}, and by that defeats the purpose of why we thought of this approach in the first place. 
One way to counter the variance is to use an effective baseline $b_\tau$.
Peters et al. \cite{4867} find that an estimate of the optimal baseline can be calculated by
\begin{equation}
	b_\tau = \frac
		{\left\langle 
			\left(\sum_{t=0}^T \nabla_\theta \log\pi(a_t|s_t) \right)^2 \sum_{t'=0}^T a_{t'} r_{t'} 
		\right\rangle}
		{\left\langle
			\left(\sum_{t=0}^T \nabla_\theta \log\pi(a_t|s_t) \right)^2
		\right\rangle},
\end{equation}
which does not affect the unbiasedness of the estimate.

@@

\begin{figure}
  \includegraphics[width=0.5\textwidth]{nat-grad-adv}
  \caption{An experiment showing where the natural gradient has a great advantage \cite{Peters_IICHR_2003}. }\label{fig:nat-grad-adv}
\end{figure}

Using the Fisher information matrix $F_\theta$, and the gradient estimate we discussed in section \ref{sec:pge} gives us the definition
\begin{equation}
  \widetilde{\nabla}_\theta J(\theta) := F^{-1}_\theta \nabla_\theta J(\theta)
  \label{eqn:nat-grad}
\end{equation}
of the natural gradient. 
The Fisher information matrix represents the certainty we have on our estimate of the gradient and is defined as the covariance of the log likelihood function of a trajectory $\tau_{\pi}^T$, which as \cite{4863} show can be written as 
\begin{equation}
  F_\theta = \int_\mathbb{S} d^\pi(s) \int_\mathbb{A} \pi(a|s) \nabla_\theta \log{\pi(a|s)} \nabla_\theta \log{\pi(a|s)}^T dads.
  \label{eqn:F}
\end{equation}
Using a value function estimator and calculating the natural gradient, we get the natural policy gradient algorithm (NPG) \cite{NIPS2017_7233}.
But, if we recall the definition \eqref{eqn:like-grad} of likelihood-ratio gradients, and the compatible function approximator from \eqref{eqn:compatible-approximator}, we get
\begin{equation}
  \nabla_\theta J(\theta) = F_\theta w.
  \label{eqn:J-equals-F}
\end{equation}
